---
title: "Powerlifting Through the ages"
author: "David Ing"
date: "April 21, 2019"
output: html_document
---


#Loading Packages and Dataset
```{r echo = FALSE}
# Load Packages
library(tidyverse)
library(lubridate)
library(DT)
library(ggplot2)
library(grid)
library(ggpubr)
library(gridExtra)

#read the data

meets = read_csv("C:/Users/David Ing/Documents/datasets/powerlifting-database/meets.csv")
openpowerlifting = read_csv("C:/Users/David Ing/Documents/datasets/powerlifting-database/openpowerlifting.csv")

```


#Data Wrangling
```{r}
Powerliftingstats = subset(openpowerlifting, 
                           select = c('Age', 'Sex', 'BestBenchKg', 'BestSquatKg', 'BestDeadliftKg', 'TotalKg'), 
                           drop = 'BestDeadliftKg' > 0, 'Age' > 0, 'BestSquatKg' > 0, 'BestBenchKg' > 0, 'TotalKg' > 0)
Powerliftingstatsclean = na.omit(Powerliftingstats[1:20000,])
Powerliftingstatsclean


clean_lifting_data = openpowerlifting %>%
  select(Name, Sex, Age, BodyweightKg,BestSquatKg, BestBenchKg, BestDeadliftKg, TotalKg, Wilks, Equipment)
clean_lifting_data = clean_lifting_data[complete.cases(clean_lifting_data),]
```


#Deadlift between Male and Females
```{r}
#Both Gender
Deadlift = ggplot(Powerliftingstatsclean, aes(x=Age, y=BestDeadliftKg)) + 
  geom_point(color = "dodgerblue3", alpha = 0.3) + 
  geom_smooth(method = 'loess', se=FALSE, color= "grey2") + 
  scale_x_continuous(limits = c(0,100),breaks = c(10,20,30,40,50,60,70,80,90,100))+ 
  scale_y_continuous(limits = c(0,400)) + 
  labs(title = "Age vs Deadlift")
Deadlift

#Female deadlift
FemaleDeadlift = ggplot(filter(Powerliftingstatsclean, Sex == 'F'), aes(x=Age, y=BestDeadliftKg))  + 
  geom_point(color = "dodgerblue3", alpha = 0.3) + 
  geom_smooth(method = 'loess', se=FALSE, color = "gray2") + 
  scale_x_continuous(limits = c(0,80),breaks = c(10,20,30,40,50,60,70,80))+ 
  scale_y_continuous(limits = c(0,300)) + 
  ggtitle("Female: Age vs Deadlift")
FemaleDeadlift


#Male Deadlift 
MaleDeadlift=ggplot(filter(Powerliftingstatsclean, Sex == 'M'), aes(x=Age, y=BestDeadliftKg))  + 
  geom_point(color = "dodgerblue3", alpha = 0.3) + 
  geom_smooth(method = 'loess', se=FALSE, color = "gray2") + 
  scale_x_continuous(limits = c(0,85),breaks = c(10,20,30,40,50,60,70,80,90))+ 
  scale_y_continuous(limits = c(0,400)) + 
  ggtitle("Male: Age vs Deadlift")                         
MaleDeadlift
```

#Bench Press Between Male and Female
```{r}
#Bench Press
Bench = ggplot(Powerliftingstatsclean, aes(x=Age, y=BestBenchKg)) + 
  geom_point(color = "red2", alpha = 0.3) +
  geom_smooth(method = 'loess', se=FALSE, color = "gray2") + 
  scale_x_continuous(limits = c(0,100),breaks = c(10,20,30,40,50,60,70,80,90,100))+ 
  scale_y_continuous(limits = c(0,500)) + 
  ggtitle("Age vs Bench")
Bench

#Female Bench
FemaleBench = ggplot(filter(Powerliftingstatsclean, Sex == 'F'), aes(x=Age, y=BestBenchKg)) + 
  geom_point(color = "red2", alpha = 0.3) + 
  geom_smooth(method = 'loess', se=FALSE, color = "gray2") + 
  scale_x_continuous(limits = c(0,80),breaks = c(10,20,30,40,50,60,70,80))+ 
  scale_y_continuous(limits = c(0,300)) +
  ggtitle("Female: Age vs Benchpress")
FemaleBench

#Male Bench 
MaleBench = ggplot(filter(Powerliftingstatsclean, Sex == 'M'), aes(x=Age, y=BestBenchKg)) + 
  geom_point(color = "red2", alpha = 0.3) + 
  geom_smooth(method = 'loess', se=FALSE, color = "gray2") + 
  scale_x_continuous(limits = c(0,85),breaks = c(10,20,30,40,50,60,70,80))+ 
  scale_y_continuous(limits = c(0,500)) +
  ggtitle("Male: Age vs Benchpress")
MaleBench
```

#Squat Between Male and Female
```{r}
#Squat
Squat = ggplot(Powerliftingstatsclean, aes(x=Age, y=BestSquatKg)) + 
  geom_point(color = "magenta3", alpha = 0.3) + 
  geom_smooth(method = 'loess', se=FALSE, color = "gray2") + 
  scale_x_continuous(limits = c(0,100),breaks = c(10,20,30,40,50,60,70,80,90,100))+ 
  scale_y_continuous(limits = c(0,500)) + 
  ggtitle("Age vs Squat")
Squat

#Female Squat
FemaleSquat = ggplot(filter(Powerliftingstatsclean, Sex == 'F'), aes(x=Age, y=BestSquatKg)) + 
  geom_point(color = "magenta3", alpha = 0.3) + 
  geom_smooth(method = 'loess', se=FALSE, color = "gray2") + 
  scale_x_continuous(limits = c(0,80),breaks = c(10,20,30,40,50,60,70,80))+ 
  scale_y_continuous(limits = c(0,500)) +
  ggtitle("Female: Age vs Squat")
FemaleSquat

#Male Squat
MaleSquat = ggplot(filter(Powerliftingstatsclean, Sex == 'M'), aes(x=Age, y=BestSquatKg)) + 
  geom_point(color = "magenta3", alpha = 0.3) + 
  geom_smooth(method = 'loess', se=FALSE, color = "gray2") + 
  scale_x_continuous(limits = c(0,85),breaks = c(10,20,30,40,50,60,70,80))+ 
  scale_y_continuous(limits = c(0,500)) +
  ggtitle("Male: Age vs Squat")
MaleSquat
```

#Total Kg
```{r}
#Total Kg
TotalKg = ggplot(Powerliftingstatsclean, aes(x=Age, y=TotalKg)) + 
  geom_point(color = "Green4", alpha = 0.3) + 
  geom_smooth(method = 'loess', se=FALSE, color = "gray2") + 
  scale_x_continuous(limits = c(0,85),breaks = c(10,20,30,40,50,60,70,80))+ 
  scale_y_continuous(limits = c(0,500)) +
  ggtitle("Age vs TotalKg")
TotalKg

#Men Total Kg
MaleTotalKg = ggplot(filter(Powerliftingstatsclean, Sex == 'M'), aes(x=Age, y=TotalKg)) + 
  geom_point(color = "Green4", alpha = 0.3) + 
  geom_smooth(method = 'loess', se=FALSE, color = "gray2") + 
  scale_x_continuous(limits = c(0,90),breaks = c(10,20,30,40,50,60,70,80,90))+ 
  scale_y_continuous(limits = c(0,500)) +
  ggtitle("Male: Age vs TotalKg")
MaleTotalKg

#Female TotalKg
FemaleTotalKg = ggplot(filter(Powerliftingstatsclean, Sex == 'F'), aes(x=Age, y=TotalKg)) + 
  geom_point(color = "Green4", alpha = 0.3) + 
  geom_smooth(method = 'loess', se=FALSE, color = "gray2") + 
  scale_x_continuous(limits = c(0,80),breaks = c(10,20,30,40,50,60,70,80))+ 
  scale_y_continuous(limits = c(0,500)) +
  ggtitle("Female: Age vs TotalKg")
FemaleTotalKg

```

#Comparison Between Male and Female
```{r}
MaleComparisonPage = ggarrange(MaleBench, MaleDeadlift, MaleSquat, MaleTotalKg)
MaleComparisonPage
FemaleComparisonPage = ggarrange(FemaleBench, FemaleDeadlift, FemaleSquat, FemaleTotalKg)
FemaleComparisonPage

```

#Maintaining Strength
```{r}
strength_dataset = clean_lifting_data %>%
  group_by(Name, Sex) %>%
  summarise(MaxSquat = max(BestSquatKg), MinSquat = min(BestSquatKg),
            MaxBench = max(BestBenchKg), MinBench = min(BestBenchKg),
            MaxDeadlift = max(BestDeadliftKg), MinDeadlift = min(BestDeadliftKg),
            BestTotalKg = max(TotalKg), LowTotalKg = min(TotalKg),
            BestWilks = max(Wilks), LowWilks = min(Wilks))

strength_drop_dataset = strength_dataset %>%
  mutate(squat_drop = (MaxSquat - MinSquat)/MaxSquat,
         bench_drop = (MaxBench - MinBench)/MaxBench,
         deadlift_drop = (MaxDeadlift - MinDeadlift)/MaxDeadlift,
         totalkg_drop = (BestTotalKg - LowTotalKg)/BestTotalKg,
         wilks_drop = (BestWilks - LowWilks)/BestWilks) %>%
  select(Name,Sex,squat_drop,bench_drop,deadlift_drop,totalkg_drop,wilks_drop)%>%
  filter(squat_drop !=0 & bench_drop!= 0 & deadlift_drop != 0 & totalkg_drop != 0 & wilks_drop != 0)%>%
  mutate_at(vars(squat_drop,bench_drop,deadlift_drop,totalkg_drop,wilks_drop),
            funs(round(.,2)))

drop_gender = strength_drop_dataset %>%
  group_by(Sex) %>%
  summarise_at(vars(squat_drop,bench_drop,deadlift_drop,totalkg_drop,wilks_drop), mean)

```

#Multiple Linear Regression
```{r}
linear_lifting_set = openpowerlifting %>%
  select(Sex, Equipment, Age, BodyweightKg,BestSquatKg, BestBenchKg, BestDeadliftKg,Wilks)
linear_lifting_set = linear_lifting_set[complete.cases(linear_lifting_set),]

linear_lifting_set$Sex <- factor(linear_lifting_set$Sex)
 
fit_lift <- lm(Wilks ~ Sex + Equipment + Age + BodyweightKg +
               BestSquatKg+ BestBenchKg + BestDeadliftKg, data = linear_lifting_set)



summary(fit_lift)
```
R^2 is .9239 which means that 92.7 % of the toal variation can be explained by its linear relationship with its predictors. Since the p-value is less than 0.5, the linear relationship between the response variable and predictors is significant.

#Full Model
```{R}
fit_lift_full = lm(Wilks ~.^2, data = linear_lifting_set)
summary(fit_lift_full)


```
Since p-value is for the f-test is less than 0.05, We can reject the null hypothesis. Therefore, we prefer the model with interactive terms.

Selecting Model
```{r}
anova(fit_lift,fit_lift_full)

```
From using the anova function, we can see the p-value of the model with interaction is lower than the with no interaction. It's also lower than 0.06. Therefore, we can conclude that the full model does a better job with fitting the data.

Stepwise Regression
```{r}
fit.backward <- step(fit_lift_full, direction = "backward")
fit.forward <- step(fit_lift_full, direction = "forward")
fit.both <- step(fit_lift_full, direction = "both")
```
THe lowest AIC will be chosen as the best model. The lowest model had a AIC of 334140.8

Testing out best Model
```{r}
best_wilks_model <- lm(Wilks~(Sex + Equipment + Age + BodyweightKg + BestSquatKg + 
    BestBenchKg + BestDeadliftKg)^2, data = linear_lifting_set)

summary(best_wilks_model)

```

#Daignostics

Residual Plot
```{r}
resd <- residuals(best_wilks_model)
pred <- fitted(best_wilks_model)
plot(pred,resd, xlab = "Predicted Values", ylab = "Residuals", main = "Residual Versus Predicted Values")
abline(h = 0)

```

QQplot
```{r}
qqnorm(resd, pch = )
qqline(resd)


```
THe qqplot look right-skewed

Histogram
```{r}
hist(resd, main = "Histogram of Residual", xlim = c(-45,70))

```

#Scatter Plot of Residuals
```{r}
plot(resd, main = "Scatter plot of residuals", pch = 19)


```


